Poster Data Analysis

### 2018-09-07 notebooks/poster_analysis.Rmd
updated figures and included heatmap and mic-plot

```{r, include = FALSE}
# Libraries
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr, ggplot2, tidyr)

@@

## COnfidence intervals
get_binCI <- function(x, n) as.numeric(setNames(binom.test(x,n)$conf.int*100,
                                                c("lwr", "upr")))

# Function that returns a filtered dataframe based on the string "genes"
filter_mut_table <- function(df) {
  report_genes <- unique(df$gene)
  grep_genes <- c()
  for (gene in genes) {
    for (g in report_genes) {
      if (grepl(gene, g, ignore.case = T) == TRUE) {
        grep_genes <- c(grep_genes, g)
      }
    }
  }
  df <- df %>%
    filter(gene %in% grep_genes)
  return(df)
}

# Function that returns a filtered dataframe based on the string "acquired_genes"
filter_acquired_table <- function(df) {
  report_genes <- unique(df$gene)
  grep_genes <- c()
  for (gene in acquired_genes) {
    for (g in report_genes) {
      if (grepl(gene, g, ignore.case = T) == TRUE) {
        grep_genes <- c(grep_genes, g)
      }
    }
  }
  df <- df %>%
    filter(gene %in% grep_genes) %>%
    mutate(gene = gsub("_", "", gene))
  return(df)
}

```

@@

total_data <- isolate_data %>%
  left_join(mic_data) %>%
  left_join(mlst_results) %>%
  rename("m_gyrB" = gyrB) %>%
  left_join(resfinder_results) %>%
  select(-type) %>%
  left_join(megares_results) %>%
  select(-type)
```

@@

resfinder_results %>%
  left_join(isolate_data, by = "id") %>%
  gather(key, value, gene_names) %>%
  group_by(key, value, species) %>%
  count() %>%
  ungroup() %>%
  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
  spread(value, n, fill = 0) %>%
  rowwise() %>%
  mutate(Total = Present + Absent,
         Percent = round(Present/Total*100, 1),
         lwr = round(get_binCI(Present, Total)[1], 1),
         upr = round(get_binCI(Present, Total)[2], 1)) %>%
  ggplot(aes(key, Percent, fill = species))+
  geom_col(color = "black",
           position = position_dodge(1))+
  geom_errorbar(aes(ymin = lwr, ymax = upr),
                width = 0.6,
                position = position_dodge(1))+
  labs(y = "Percent (%) of isolates")+
  guides(fill = FALSE)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, size = 8, hjust = 1, vjust = 0.4),
        axis.title.x = element_blank())+
  facet_wrap(~species)
```

@@

megares_results %>%
  left_join(isolate_data, by = "id") %>%
  gather(key, value, mut_names) %>%
  group_by(key, value, species) %>%
  count() %>%
  ungroup() %>%
  mutate(value = if_else(value == 1, "Mutant", "Wild Type"),
         compl = complete.cases(.)) %>%
  filter(compl == TRUE) %>%
  spread(value, n, fill = 0) %>%
  rowwise() %>%
  mutate(Total = Mutant + `Wild Type`,
         Percent = round(Mutant/Total*100, 1),
         lwr = round(get_binCI(Mutant, Total)[1], 1),
         upr = round(get_binCI(Mutant, Total)[2], 1)) %>%
  ggplot(aes(key, Percent, fill = species))+
  geom_col(color = "black",
           position = position_dodge(1))+
  geom_errorbar(aes(ymin = lwr, ymax = upr),
                width = 0.6,
                position = position_dodge(1))+
  labs(y = "Percent (%) of isolates")+
  guides(fill = FALSE)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, size = 8, hjust = 1, vjust = 0.4),
        axis.title.x = element_blank())+
  facet_wrap(~species)
```

## MIC plots
```{r}

```

>>>>

```{r, include = FALSE}
# Libraries
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats)

@@

## COnfidence intervals
get_binCI <- function(x, n) as.numeric(setNames(binom.test(x,n)$conf.int*100,
                                                c("lwr", "upr")))
```

@@

total_data <- isolate_data %>%
  left_join(mic_data, by = "id") %>%
  left_join(mlst_results, by = "id") %>%
  rename("m_gyrB" = gyrB) %>%
  left_join(resfinder_results, by = "id") %>%
  select(-type) %>%
  left_join(megares_results, by = "id") %>%
  select(-type)
```

@@

resfinder_results %>%
  left_join(isolate_data, by = "id") %>%
  gather(key, value, gene_names) %>%
  group_by(key, value, species) %>%
  count() %>%
  ungroup() %>%
  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
  spread(value, n, fill = 0) %>%
  rowwise() %>%
  mutate(Total = Present + Absent,
         Percent = round(Present/Total*100, 1),
         lwr = round(get_binCI(Present, Total)[1], 1),
         upr = round(get_binCI(Present, Total)[2], 1)) %>%
  ggplot(aes(key, Percent, fill = species))+
  geom_col(color = "black",
           position = position_dodge(1))+
  geom_errorbar(aes(ymin = lwr, ymax = upr),
                width = 0.6,
                position = position_dodge(1))+
  scale_fill_brewer(palette = "Paired")+
  labs(y = "Percent (%) of isolates")+
  guides(fill = FALSE)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, size = 8, hjust = 1, vjust = 0.4),
        axis.title.x = element_blank())+
  facet_wrap(~species)
```

@@

megares_results %>%
  left_join(isolate_data, by = "id") %>%
  gather(key, value, mut_names) %>%
  group_by(key, value, species) %>%
  count() %>%
  ungroup() %>%
  mutate(value = if_else(value == 1, "Mutant", "Wild Type"),
         compl = complete.cases(.)) %>%
  filter(compl == TRUE) %>%
  spread(value, n, fill = 0) %>%
  rowwise() %>%
  mutate(Total = Mutant + `Wild Type`,
         Percent = round(Mutant/Total*100, 1),
         lwr = round(get_binCI(Mutant, Total)[1], 1),
         upr = round(get_binCI(Mutant, Total)[2], 1)) %>%
  ggplot(aes(key, Percent, fill = species))+
  geom_col(color = "black",
           position = position_dodge(1))+
  geom_errorbar(aes(ymin = lwr, ymax = upr),
                width = 0.6,
                position = position_dodge(1))+
  scale_fill_brewer(palette = "Paired")+
  labs(y = "Percent (%) of isolates")+
  guides(fill = FALSE)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, size = 8, hjust = 1, vjust = 0.4),
        axis.title.x = element_blank())+
  facet_wrap(~species)
```

## MIC plot

```{r message=FALSE, warning=FALSE}
selected_genes <- c("qnrA1","qnrB19","qnrB6","qnrB60","qnrS1","qnrS2","qnrS4")
selected_mut <- c("gyrA","gyrB","parC","parE","marR")

micplot_data <- mic_data %>%
  select(id, CIP) %>%
  left_join(megares_results, by = "id") %>%
  left_join(resfinder_results, by = "id") %>%
  select(-c(acrB, type.x, type.y)) %>%
  mutate(compl = complete.cases(.)) %>%
  filter(compl == TRUE) %>%
  select_at(vars(id, CIP,selected_mut, selected_genes)) %>%
  mutate(none = if_else(rowSums(.[,3:ncol(.)]) != 0, 0, 1)) %>%
  group_by_at(vars(selected_mut, selected_genes, none)) %>%
  { mutate(ungroup(.), group = group_indices(.)) } %>%
  gather(gene, value, -c(id, CIP, group)) %>%
  mutate(gene = ifelse(value == 0 & gene != "none", NA, gene),
         value = ifelse(gene == "none" & value == 0, NA, value)) %>%
  na.omit()

factor_breaks <- c(0.015, 0.03, 0.06, 0.12, 0.25, 0.5, 1, 2, 4, 8, 16)

factor_levels <- c("0.015", "0.03", "0.06", "0.12", "0.25", "0.5", "1", "2", "4", 
                   "8", "16")

p0 <- micplot_data %>%
  group_by(id, group) %>%
  count() %>%
  group_by(group) %>%
  count() %>%
  ggplot(aes(factor(group), nn, fill = factor(group)))+
  geom_col(color = "black")+
  guides(fill = FALSE)+
  scale_fill_viridis(discrete = TRUE,
                     direction = -1)+
  theme_light()+
  theme(axis.title = element_blank(),
        axis.text.y = element_text(size = 8),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid = element_blank())

p1 <- micplot_data %>%
  group_by(CIP, group, gene) %>%
  count() %>%
  ggplot(aes(factor(group), CIP, fill = factor(group), size = n))+
  geom_point(pch = 21)+
  scale_y_continuous(trans = "log10",
                     labels = factor_levels,
                     breaks = factor_breaks)+
  guides(fill = FALSE,
         size = FALSE)+
  scale_fill_viridis(discrete = TRUE,
                     direction = -1)+
  theme_light()+
  theme(axis.title = element_blank(),
        axis.text.y = element_text(size = 8),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid = element_blank())

p2 <- ggplot(micplot_data,
       aes(factor(group),
           factor(gene),
           fill = factor(group)))+
  geom_point(shape = 23,
             pch = 21,
             size = 2)+
  geom_hline(yintercept = 5.5)+
  guides(fill = FALSE)+
  scale_fill_viridis(discrete = TRUE,
                     direction = -1)+
  labs(y = "Genes")+
  theme_light()+
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 8),
        axis.ticks.x = element_blank(),
        panel.grid = element_blank())

plot_grid(p0, p1, p2, align = "v", nrow = 3, rel_heights = c(1/4, 2/4, 1/4))
```

## Heatmap plot

```{r, fig.width = 10, fig.height=30}
species_data <- select(isolate_data, c(id, species))

megares_long <- megares_results %>%
  gather(gene, value, -c(id, type)) %>%
  filter(is.na(value) == FALSE)

resfinder_long <- resfinder_results %>%
  gather(gene, value, -c(id, type))

megares_long %>%
  rbind(resfinder_long) %>%
  mutate(type2 = ifelse(type == "mut", 1, 2)) %>%
  spread(id, value, fill = 0) %>%
  gather(key = "id", value = "result", -c(gene, type, type2)) %>%
  left_join(species_data, by = "id") %>%
  ggplot(aes(fct_reorder(gene, type2),
             factor(id,
                    levels = unique(.[order(.$species,
                                                          .$id),
                                                    "id",
                                                    drop = TRUE])),
             fill = species,
             alpha = result)) +
  geom_tile(color = "white")+
  geom_vline(xintercept = 14.5)+
  scale_fill_brewer(palette = "Paired")+
  theme_minimal()+
  labs(fill = NULL)+
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.3,
                                   size = 7),
        axis.title = element_blank(),
        axis.text.y = element_blank(),
        panel.grid = element_blank(),
        legend.position = "right")+
  guides(alpha = FALSE)+
  coord_fixed()
```
